%%       Make List of References
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: I use BibTeX, but have coded in thebibliography environment
%%       to use for this example only.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


									% unstr is the same as
\begin{thebibliography}{1}
\bibliographystyle{plain}


\bibitem{sperber} A. Adolphson and S. Sperber. 
\newblock $p$-adic Estimates for Exponential Sums and the of Chevalley-Warning.
\newblock {\it Ann. Sci. Ec. Norm. Super.}, $4^{e}$ s\'erie,  {\bf 20}, 545--556, 1987.

\bibitem{acgmr} R. A. Arce-Nazario, F. N. Castro, O. E. Gonz\'alez, L. A. Medina and I. M. Rubio.
\newblock New families of balanced symmetric functions and a generalization of Cuscik, Li and P. St$\check{\mbox{a}}$nic$\check{\mbox{a}}$.
\newblock {\it Designs, Codes and Cryptography} {\bf 86}, 693--701, 2018.

\bibitem{ax} J. Ax.
\newblock Zeros of polynomials over finite fields. 
\newblock {\it Amer. J. Math.}, {\bf 86}, 255--261, 1964.

\bibitem{BCP} M. L. Bileschi, T.W. Cusick and D. Padgett.
\newblock Weights of Boolean cubic monomial rotation symmetric functions.
\newblock {\it Cryptogr. Commun.}, {\bf 4}, 105--130, 2012.

\bibitem{browncusick} A. Brown and T. W. Cusick.
\newblock Recursive weights for some Boolean functions. 
\newblock  {\it J. Math. Cryptology}, {\bf 6(2)}, 105--135, 2012.

\bibitem{cai} J. Cai, F. Green and T. Thierauf. 
\newblock On the correlation of symmetric functions.
\newblock {\it Math. Systems Theory}, {\bf 29}, 245--258, 1996.

\bibitem{canteaut} A. Canteaut and M. Videau.
\newblock Symmetric Boolean Functions.
\newblock {\it IEEE Trans. Inf. Theory} {\bf 51(8)}, 2791--2881, 2005.

\bibitem{Davis}
\newblock Philip Davis. Circulant Matrices.
\newblock Chelsea publishing, Second Edition,1994.

\bibitem{cgm3} F. N. Castro, O. E. Gonz\'alez and L. A. Medina.
\newblock Diophantine Equations With Binomial Coefficients and Perturbations of Symmetric Boolean Functions.
\newblock {\it IEEE Trans. Inf. Theory}, {\bf 64(2)}, 1347--1360, 2018.

\bibitem{cm1} F. N. Castro and L. A. Medina. 
\newblock Linear Recurrences and Asymptotic Behavior of Exponential Sums of Symmetric Boolean Functions. 
\newblock {\it Elec. J. Combinatorics}, 18:\#P8, 2011.

\bibitem{cm2} F. N. Castro and L. A. Medina. 
\newblock Asymptotic Behavior of Perturbations of Symmetric Functions.  
\newblock {\it Annals of Combinatorics}, 18:397--417, 2014.

\bibitem{cm3} F. N. Castro and L. A. Medina. 
\newblock Modular periodicity of exponential sums of symmetric Boolean functions.
\newblock {\it Discrete Appl. Math.} {\bf 217}, 455--473, 2017.

\bibitem{cms} F. N. Castro, L. A. Medina and P. St$\check{\mbox{a}}$nic$\check{\mbox{a}}$.
\newblock Generalized Walsh transforms of symmetric and rotation symmetric Boolean functions are linear recurrent.
\newblock {\it Appl. Algebra Eng. Commun. Comput.}, DOI 10.1007/s00200-018-0351-5, 2018.

\bibitem{ccms} F. N. Castro, R. Chapman, L. A. Medina, and L. B. Sep\'ulveda.  
\newblock Recursions associated to trapezoid, symmetric and rotation symmetric functions over Galois fields.
\newblock {\it Discrete Mathematics.},{\bf 341}, 1915--1931, 2018.

\bibitem{cusick4} T. W. Cusick. 
\newblock Hamming weights of symmetric Boolean functions.
\newblock {\it Discrete Appl. Math.} {\bf 215}, 14--19, 2016.

\bibitem{cusickArXiv} T. W. Cusick.
\newblock Weight recursions for any rotation symmetric Boolean functions.
\newblock {\it IEEE Trans. Inf. Theory}, {\bf 64}, 2962 - 2968, 2018.

\bibitem{cusickjohns} T. W. Cusick and B. Johns.
\newblock Recursion orders for weights of Boolean cubic rotation symmetric functions.
\newblock {\it Discr. Appl. Math.}, {\bf 186}, 1--6, 2015.

\bibitem{cusick2} T. W. Cusick, Y. Li, and  P. St$\check{\mbox{a}}$nic$\check{\mbox{a}}$.
\newblock Balanced Symmetric Functions over $GF(p)$.
\newblock {\it IEEE Trans. Inf. Theory}, {\bf 5}, 1304--1307, 2008.

\bibitem{cusickstanica} T.W. Cusick and P. St$\check{\mbox{a}}$nic$\check{\mbox{a}}$.
\newblock Fast evaluation, weights and nonlinearity of rotation symmetric functions.
\newblock {\it Discr. Math.}, {\bf 258}, 289--301, 2002.

\bibitem{dalaimaitrasarkar} D. K. Dalai, S. Maitra and S. Sarkar. 
\newblock Results on rotation symmetric Bent functions.
\newblock {\it Second International Workshop on Boolean Functions: Cryptography and
Applications, BFCA'06}, publications of the universities of Rouen and Havre, 137--156, 2006.

\bibitem{fengliu} K. Feng and F. Liu. 
\newblock New Results On The Nonexistence of Generalized Bent Functions. 
\newblock {\it IEEE Trans. Inf. Theory} {\bf 49},  3066--3071, 2003.

\bibitem{ggz} Y. Guo, G. Gao, Y. Zhao.
\newblock Recent Results on Balanced Symmetric Boolean Functions.
\newblock {\it IEEE Trans. Inf. Theory} {\bf 62 (9)}, 5199--5203, 2016.

\bibitem{hell} M. Hell, A. Maximov and S. Maitra. 
\newblock On efficient implementation of search strategy for rotation symmetric Boolean functions. 
\newblock {\it Ninth International Workshop on Algebraic and Combinatorial Coding Theory, ACCT 2004}, Black Sea Coast, Bulgaria,
2004.

\bibitem{hg} Y. Hu and G. Xiao.
\newblock Resilient Functions Over Finite Fields.
\newblock {\it IEEE Trans. Inf. Theory} {\bf 49},  2040--2046, 2003.

\bibitem{ims} E. J. Iona\c{s}cu, Thor Martinsen, Pantelimon St$\check{\mbox{a}}$nic$\check{\mbox{a}}$.
\newblock Bisecting binomial coefficients.
\newblock {\it Discrete Appl. Math.} {\bf 227} (2017) 70--83.

\bibitem{fspectrum} M. Kolountzakis, R. J. Lipton, E. Markakis, A. Metha and N. K. Vishnoi. 
\newblock On the Fourier Spectrum of Symmetric Boolean Functions.
\newblock {\it Combinatorica}, {\bf  29}, 363--387, 2009.

\bibitem{KScW} P.V. Kumar, R.A. Scholtz, and L.R. Welch. 
\newblock Generalized Bent Functions and Their Properties.
\newblock {\it J. Combinatorial Theory (A)}, {\bf 40}, 90--107, 1985.

\bibitem {licusick1} Y. Li and T.W. Cusick. 
\newblock Linear Structures of Symmetric Functions over Finite Fields.
\newblock {Inf. Processing Letters} {\bf 97}, 124--127, 2006.

\bibitem{licusick2} Y. Li and T. W. Cusick. 
\newblock Strict Avalanche Criterion Over Finite Fields.
\newblock {\it J. Math. Cryptology}  {\bf 1(1)}, 65--78, 2006.

\bibitem{llm}M. Liu, P. Lu and G.L. Mullen. 
\newblock Correlation-Immune Functions over Finite Fields.
\newblock {\it IEEE Trans. Inf. Theory} {\bf 44}, 1273--1276, 1998.

\bibitem{maxhellmaitra} A. Maximov, M. Hell and S. Maitra. 
\newblock Plateaued Rotation Symmetric Boolean Functions on Odd Number of Variables. 
\newblock {\it First Workshop on Boolean Functions:Cryptography and Applications, BFCA'05}, publications of the universities of Rouen and Havre, 83--104, 2005.

\bibitem{mitchell}  C. Mitchell.
\newblock Enumerating Boolean functions of cryptographic significance.
\newblock {\em J. Cryptology} {\bf 2} (3), 155--170, 1990.

\bibitem{mm1} O. Moreno and C. J. Moreno.
\newblock Improvement of the Chevalley-Warning and the Ax-Katz theorems.
\newblock {\it Amer. J. Math.}, {\bf 117}, 241--244, 1995.

\bibitem{mm} O. Moreno and C. J. Moreno.
\newblock The MacWilliams-Sloane Conjecture on the Tightness of the Carlitz-Uchiyama Bound and the Weights of Dual of BCH Codes.
\newblock {\it IEEE Trans. Inform. Theory}, {\bf 40},  1894--1907, 1994.

\bibitem{parkerpott} M. G. Parker and A. Pott. 
\newblock On Boolean functions which are bent and negabent.
\newblock {\it Proc. Int. Conf. Sequences, Subsequences, Consequences},
LNCS-4893, 9--23, 2007.

\bibitem{piequ} J. Pieprzyk and C.X. Qu.
\newblock Fast hashing and rotation-symmetric functions.
\newblock {\it J. Universal Comput. Sci.}, {\bf 5 (1)}, 20--31, 1999.

\bibitem{rp} C. Riera and M. G. Parker.
\newblock Generalized bent criteria for Boolean functions.
\newblock {\it IEEE Trans. Inform. Theory} {\bf 52} (9), 4142--4159, 2006.

\bibitem{fdegree} A. Shpilka and A. Tal.
\newblock On the Minimal Fourier Degree of Symmetric Boolean Functions.
\newblock {\it Combinatorica}, {\bf 88}, 359--377, 2014.

\bibitem{stanicamaitra} P. St$\check{\mbox{a}}$nic$\check{\mbox{a}}$ and S. Maitra. 
\newblock Rotation Symmetric Boolean Functions -- Count and Cryptographic Properties. 
\newblock {\it Discr. Appl. Math.}, {\bf 156}, 1567--1580, 2008

\bibitem{stanicamaitraclark} P. St$\check{\mbox{a}}$nic$\check{\mbox{a}}$, S. Maitra and J. Clark. 
\newblock Results on Rotation Symmetric Bent and Correlation Immune Boolean Functions. 
\newblock {\it Fast Software Encryption, FSE 2004}, Lecture Notes in Computer Science, {\bf 3017}, 161--177. SpringerVerlag, 2004.

\end{thebibliography}

\addcontentsline{toc}{extrachapter}{BIBLIOGRAPHY}%







%\bibliography{referencefiles/references.bib} 		% Use the References.bib file

%\backmatter %

%\biography{										% Add the Biography pages if you want.
%\input{backfiles/Biography.tex}			
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  Create a General Audience Abstract      %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{simpleenv}{}{}{}{}%
%\pagestyle{empty}
%\begin{flushleft}
%\Title\\*[\BaseDiff\baselineskip]
%\FullName\\%
%(787) XXX-XXXX\\%
%Department of \Department \\%
%Chair: \Chair\\%
%Degree: \DegreeType\\%
%Graduation Date: \GradMonth~\GradYear%
%\end{flushleft}
%\GoDouble%
%\input{backfiles/GeneralAudienceAbstract.tex}
%\end{simpleenv}

